Hyperbolicity of Augmented Links in the Thickened Torus

### 2021-08-21 main.tex
fixed up the torihedral decomposition to make half-twist stuff clearer

Given a twist reduced diagram of a link $K$, \emph{augmenting} is a process in
which an unknotted circle component (augmentation) is added to one or more twist
regions (a single crossing or a maximal string of bigons) of $K$. The added circle
component allows us to remove full twists at the twist region of $K$. If the
twist region has an odd number of crossings then all but one crossing is
removed, whereas if the twist region has an even number of crossings then all
are removed. We can remove full twists by a standard argument using a Dehn twist 
on the complement of the crossing circle. The newly obtained link is called an 
\emph{augmented link} and the newly obtained diagram is called an 
\emph{augmented link diagram}. See Figure
\ref{fig:Augmentations}.

@@

Menasco \cite{Menasco} showed there are decompositions of the complements of
alternating links in $\Sp^3$ into two topological polyhedra, a top polyhedron and a bottom
polyhedron.  For alternating links $K$ in the thickened torus, Champanerkar,
Kofman and Purcell \cite{CKP2} showed that there is a decomposition of the
complement of $K$ into objects called torihedra, which we think of as
counterparts to Menasco's decomposition of links in $\Sp^3$ into polyhedra, for
links in the thickened torus. Just like Menasco's decomposition, one obtains a
top and a bottom torihedron.

In \secref{s:auglinks} we show that for augmented
links in the thickened torus one can also obtain a decomposition of the
complement into a top and bottom torihedron.  In \secref{s:hyperbolicity}, we
prove that many augmented alternating links in the thickened torus are
hyperbolic.

@@

 \begin{figure}
 \centering  
 \begin{tabular}{cc}
 \includegraphics [height=4cm]{augmentation1.png}&
  \includegraphics [height=4cm]{augmentation2.png}\\
  (a)&(b)
  \end{tabular}
 \caption{The left shows a pretzel knot before augmentation and the right shows after augmentation}
 \label{fig:augmentationS3}
 \end{figure}

@@

Now we can define augmentation for a link in $\torus \times I$ the same way we
define augmentation for links in $\Sp^3$:
%For a link in $\torus \times I$, the
%crossing circles are added to the diagram projected onto $\torus \times \{0\}$.
%Let $D(L)$ be a twist reduced diagram of a link $L$ in $\torus \times I$, we define
%\emph{augmenting} as a process in which an unknotted circle component is added
%to one or more twist regions of $D(L)$. See Figure \ref{fig:Augmentations}

\begin{definition}
Let $D(K)$ be a twist reduced diagram of a link $K$ in $\torus \times I$.
We define
\emph{augmenting} as a process in which an unknotted circle component,
called a \emph{crossing circle},
is added to one or more twist regions of $D(K)$
(see Figure \ref{fig:Augmentations});
we call the resulting link $L$ an \emph{augmented link obtained from $K$}.
We say $L$ is \emph{fully} augmented if $L$ is obtained by augmenting
$K$ at \emph{every} crossing/twist site.
\label{d:augmentation}
\end{definition}

As pointed out in the introduction,
after augmenting a twist region,
a standard Dehn twist argument allows us to remove
a full twist (that is, two bigons).
We say an augmentation has a \emph{half twist}
if at least one of the augmented twist regions
has an odd number of bigons.

@@

\begin{definition}
Let $G$ be a graph with no self-loops on the torus.

\label{d:diagram-bowtie}
\end{definition}

@@

\begin{proof}
We will begin by assuming that $L$ has no half twists.
Let $L = K \cup C$, with $C$ being the collection of crossing circles.
%Arrange the link diagram of $L$ in the following way:
Arrange $L$ in the following way:
place the circle components in $C$ perpendicular to
the projection plane $\torus \times \{0\}$,
and leave the remaining part of the link $K \subseteq L$
lying in the projection plane (except at crossings).
We now place a \emph{crossing edge} at each crossing of the original link $K$,
connecting the top and bottom strands at the crossing
(see Figure \ref{fig:crossingArc} left).
%so that for each crossing edge,
%one end of the edge lies on a bottom strand while the other
%end lies on a top strand as in Figure \ref{fig:crossingArc} left.

\begin{figure} 
\centering 
\includegraphics[width=4cm]{crossingArc.png}
\includegraphics[width=5cm]{crossingPush.png}
\caption{Left: The black strands are
part of the link and the red strand is the crossing edge. Right: The blue and
red edges represent the split crossing edges and the shaded half disk is bounded
by the crossing circle} 
\label{fig:crossingArc} 
\end{figure}

Next we construct the graph that will define the ``top'' torihedron.
We view the link from the point at infinity from the top end
($\torus \times \{1\}$) of the thickened torus.
Push the top strand to the bottom strand,
splitting the crossing edge into two
identical edges and spreading them apart
as in Figure \ref{fig:crossingArc} right.
We push the link
components to infinity and stretch the crossing edge so that we have flattened
the link onto $\torus \times \{0\}$ except for the crossing circles,
which remains perpendicular to the projection plane.

%Now place a disk on each crossing circle, so that the disk is bounded by
%the crossing circle.
Now for each crossing circle $c$, consider a spanning disk $B_c$;
$B_c$ intersects the projection plane $\torus \times \{0\}$
along the red segments in Figure \ref{fig:falGluings},
which consists of three edges.
We then cut $\torus \times I$ along $\torus \times \{0\}$ and
focus on the top half, $\torus \times [0,1)$.
(We will follow the same method on the
bottom half to obtain the second identical torihedron.)
The spanning disks we placed for each crossing circle are now cut in half
along the red segment.
Each half of the disk is now bounded by the
projection plane and the semi-circle arc of the crossing circle.
We push down on the crossing circle and split the disk into
two identical disks.
We then push the arc of each crossing circle to infinity,
collapsing them to ideal vertices.
We obtain two triangular faces which represent the disk which look like a
bow-tie as in Figure \ref{fig:falGluings} (a).

@@

We repeat the steps for the bottom half of $\torus \times I$, $\torus \times
(-1,0]$. Then we get two torihedra. The graph of each will come from
crossing edges and edges from crossing circles.
Now, if there are half twists,
we decompose the complement of the link the same way
(as if there are no half twists)
but we identify the two bow-ties as in Figure \ref{fig:falGluings} (b).
Finally,
we obtain the complement of the link by gluing the two torihedra with the gluing
information given by identifying crossing edges and triangles of the bow-tie. We
glue the faces of the torihedra which do not correspond to a bow-tie with a
$2\pi/n$ twist where $n$ is the number of sides of each face as in Figure
\ref{fig:top-bottom} clockwise or counterclockwise.

For future reference, we will denote the graph for the top and bottom
torihedra by $\Gamma_T(L)$ and $\Gamma_B(L)$, respectively,
where both graphs are viewed from the cone point of the top torihedron
$\torus \times \{1\}$.
Note that if $L = K$ is the non-augmented link,
$\Gamma_T(L)$ is simply the link projection of $K$,
and in fact $\Gamma_T(K) = \Gamma_B(K)$.

@@

\begin{figure}[h] 
\centering 
\includegraphics[height=3cm]{fig-6.png} 
\caption{Left: The crossing arc is the edge in red.
Middle: Picture of splitting the crossing edge. Right: The link component is
pushed off to infinity.}
	\label{fig:step_three}
\end{figure}

\begin{figure}[h] 
\centering 
\includegraphics[height=3cm]{top-bottom.png} 
	\caption{Left: The top torihedron. Right: The bottom torihdron with rotation for face gluing.} 
\label{fig:top-bottom}
\end{figure}

@@

More generally, if $K$ is as above with a twist reduced diagram
containing bigons,
and $L$ is obtained by augmenting $K$
at every twist region with at least one bigon
(and possibly other crossings),
then $L$ is hyperbolic.
\end{theorem}

>>>>

Given a twist-reduced diagram of a link $K$, \emph{augmenting} is a process in
which an unknotted circle component (augmentation) is added to one or more twist
regions (a single crossing or a maximal string of bigons) of $K$.
%%By a standard argument using a Dehn twist
%%on the complement of the crossing circle,
%%the added circle component allows us to remove full twists
%%(i.e. pairs of bigons)
%%at the twist region of $K$
%%while keeping the link complement the same.
%%If the twist region has an odd number of crossings,
%%then all but one crossing is removed,
%%whereas if the twist region has an even number of crossings,
%%then all are removed.
%We can remove full twists by a standard argument using a Dehn twist 
%on the complement of the crossing circle.
The newly obtained link is called an 
\emph{augmented link} and the newly obtained diagram is called an 
\emph{augmented link diagram}. See Figure
\ref{fig:Augmentations}.

@@

Menasco \cite{Menasco} showed that there are decompositions
of the complements of alternating links in $\Sp^3$
into two topological polyhedra,
a top polyhedron and a bottom polyhedron.
For alternating links $K$ in the thickened torus,
Champanerkar, Kofman and Purcell \cite{CKP2}
showed that there is a decomposition of the
complement of $K$ into objects called torihedra, which we think of as
counterparts to Menasco's decomposition % of links in $\Sp^3$ into polyhedra,
for links in the thickened torus;
just like Menasco's decomposition, one obtains a top and a bottom torihedron.

In \secref{s:auglinks} we show that for augmented
links in the thickened torus (not necessarily fully augmented),
one can also obtain a decomposition of the
complement into a top and bottom torihedron.
In \secref{s:hyperbolicity}, we prove that
many augmented alternating links in the thickened torus are hyperbolic.

@@

\begin{figure}
\centering  
\begin{tabular}{cc}
\includegraphics[height=4cm]{augmentation1.png}
\;\;\;\;
& \includegraphics[height=4cm]{augmentation2.png}
\\
(a)
& (b)
\end{tabular}
\caption{(a) pretzel knot before augmentation
(b) pretzel knot after augmentation;
second diagram shows
removal of full twists in the augmented twist regions.}
\label{fig:augmentationS3}
\end{figure}

@@

Now we can define augmentation for a link in $\torus \times I$ the same way we
define augmentation for links in $\Sp^3$:
%For a link in $\torus \times I$, the
%crossing circles are added to the diagram projected onto $\torus \times \{0\}$.
%Let $D(L)$ be a twist-reduced diagram of a link $L$ in $\torus \times I$, we define
%\emph{augmenting} as a process in which an unknotted circle component is added
%to one or more twist regions of $D(L)$. See Figure \ref{fig:Augmentations}

\begin{definition}
Let $D(K)$ be a twist-reduced diagram of a link $K$ in $\torus \times I$.
We define
\emph{augmenting} as a process in which an unknotted circle component,
called a \emph{crossing circle},
is added to one or more twist regions of $D(K)$
(see Figure \ref{fig:Augmentations});
we call the resulting link $L$ an \emph{augmented link obtained from $K$}.
We say $L$ is \emph{fully} augmented if $L$ is obtained by augmenting
$K$ at \emph{every} crossing/twist site.
\label{d:augmentation}
\end{definition}

As pointed out in the introduction,
after augmenting a twist region,
a standard Dehn twist argument allows us to remove
a full twist (that is, two bigons).

\begin{definition}
We say an augmentation has a \emph{half twist}
if at least one of the augmented twist regions
has an odd number of bigons.
\label{d:half-twist}
\end{definition}

@@

\begin{definition}
Let $G$ be a graph with no self-loops on the torus.
TODO definition of adding a bow-tie
\label{d:diagram-bowtie}
\end{definition}

@@

\begin{proof}
We will begin by assuming that $L$ has no half twists.
Let $L = K \cup C$, with $C$ being the collection of crossing circles.
%Arrange the link diagram of $L$ in the following way:
Arrange $L$ in the following way:
place the circle components in $C$ perpendicular to
the projection plane $\torus \times \{0\}$,
and leave the remaining part of the link $K \subseteq L$
lying in the projection plane (except at crossings of $K$).
Thus, the projection of $L$ onto the projection plane
will be a diagram $D(K)$ of $K$
together with line segments corresponding to crossing circles.
(We demand that $D(K)$ be alternating and twist-reduced.)
As discussed before, we may remove full twists (i.e. pairs of bigons)
from the twist regions of $K$ that are augmented in $L$,
and since $L$ is assumed to have no half twists,
the two edges of $D(K)$ that go through a crossing circle
do not meet (see Figure \ref{fig:Augmentations} C).
From now on, we will take $K$ to be this modified link
(with full twists removed).

We now place a \emph{crossing edge} at each crossing of the original link $K$,
connecting the top and bottom strands at the crossing
(see Figure \ref{fig:crossingArc} (a)).
%so that for each crossing edge,
%one end of the edge lies on a bottom strand while the other
%end lies on a top strand as in Figure \ref{fig:crossingArc} (a).

\begin{figure} 
\centering 
\begin{tabular}{cc}
\includegraphics[width=4cm]{crossingArc.png}
& \includegraphics[width=5cm]{crossingPush.png}
\\
(a)
& (b)
\end{tabular}
\caption{(a) The black strands are
part of the link and the red strand is the crossing edge.
(b) The blue and
red edges represent the split crossing edges and the shaded half disk is bounded
by the crossing circle} 
\label{fig:crossingArc} 
\end{figure}

Next we construct the graph that will define the ``top'' torihedron.
We view the link from the point at infinity from the top end
($\torus \times \{1\}$) of the thickened torus.
At each crossing of $K$,
push the top strand towards the bottom strand,
splitting the crossing edge into two
identical edges and spreading them apart
as in Figure \ref{fig:crossingArc} (b).
%We push the link
%components to infinity and stretch the crossing edge so that we have flattened
%the link onto $\torus \times \{0\}$ except for the crossing circles,
%which remains perpendicular to the projection plane.

%Now place a disk on each crossing circle, so that the disk is bounded by
%the crossing circle.
Now for each crossing circle $c$, consider a spanning disk $B_c$;
$B_c$ intersects the projection plane $\torus \times \{0\}$
along the red segments in Figure \ref{fig:falGluings},
which consists of three edges.
We then cut $\torus \times I$ along $\torus \times \{0\}$ and
focus on the top half, $\torus \times [0,1)$.
(We will follow the same method on the
bottom half to obtain the second identical torihedron.)
The spanning disks we placed for each crossing circle are now cut in half
along the red segment.
Each half of the disk is now bounded by the
projection plane and the semi-circle arc of the crossing circle.
We push down on the crossing circle and split the half-disk into
two identical half-disks.
We then push the arc of each crossing circle to infinity,
collapsing them to ideal vertices.
We obtain two triangular faces which represent the half-disk which look like a
bow-tie as in Figure \ref{fig:falGluings} (a).
See Figures \ref{fig:step_one} to \ref{fig:top-bottom}
for an example.

@@

The crossing edges and edges from spanning disks of crossing circles
form the graph of the top ideal torihedron.
We repeat the steps for the bottom half of $\torus \times I$, $\torus \times
(-1,0]$. Thus we get two ideal torihedra.
%The graph of each will come from
%crossing edges and edges from spanning disks of crossing circles.

We obtain the complement of $L$ by gluing the two torihedra with the gluing
information given by identifying crossing edges and triangles of the bow-tie. We
glue the faces of the torihedra which do not correspond to a bow-tie with a
$2\pi/n$ twist where $n$ is the number of sides of each face as in Figure
\ref{fig:top-bottom} clockwise or counterclockwise.

Now, if $L$ has half-twists,
we apply the constructions above to $L'$,
which is the link obtained from $L$ by removing all half-twists
(replacing them with parallel strands).
The complement of $L$ is again obtained by
gluing the torihedra together all the same,
except that at the bow-ties coming from twist regions with a half-twist,
they are glued as in Figure \ref{fig:falGluings} (b).
%we decompose the complement of the link the same way
%(as if there are no half twists)
%but we identify the two bow-ties as in Figure \ref{fig:falGluings} (b).
%Finally,

For future reference, we will denote the graph for the top and bottom
torihedra by $\Gamma_T(L)$ and $\Gamma_B(L)$, respectively,
where both graphs are viewed from the cone point of the top torihedron
$\torus \times \{1\}$.
Note that if $L = K$ is the non-augmented link,
$\Gamma_T(L)$ is simply the link projection of $K$,
and in fact $\Gamma_T(K) = \Gamma_B(K)$.
\end{proof}

@@

\begin{figure}[h] 
\centering 
\includegraphics[height=3cm]{fig-6.png} 
\caption{Left: The crossing arc is the edge in red.
Middle: Picture of splitting the crossing edge.
Right: The link components are pushed off to infinity.}
	\label{fig:step_three}
\end{figure}

\begin{figure}[h] 
\centering 
\includegraphics[height=3cm]{top-bottom.png} 
	\caption{Left: The top torihedron.
	Right: The bottom torihdron with rotation indicating face gluing;
	TODO shouldn't it be rotating other way.} 
\label{fig:top-bottom}
\end{figure}

@@

More generally, if $K$ is as above with a twist-reduced diagram
containing bigons,
and $L$ is obtained by augmenting $K$
at every twist region with at least one bigon
(and possibly other crossings),
then $L$ is hyperbolic.
\end{theorem}
